---
title: "Ch4: Ames housing data"
output:
  html_notebook:
    toc: true
    toc_float: true
    theme: cerulean
---

Ultimately, we want to predict the price of a home based on various predictors in the `ames` housing data. 

```{r notebook setup}
library(tidymodels, tidyverse, patchwork)

# Load housing data
data(ames, package = "modeldata")
ames

```

Perform some EDA.

```{r ames EDA}
# Look at different binwidths for the distribution of housing price
map(c(30, 40, 50, 60), ~ ggplot(ames, aes(Sale_Price)) + 
      geom_histogram(bins = .x) + 
      geom_vline(aes(xintercept = median(Sale_Price)), 
                     linetype = "dashed", color = "red") + 
      labs(title = paste0("Bin width: ", .x))) %>% 
  wrap_plots() + 
  plot_annotation(title = "Distribution of sale price in Ames, Iowa housing data", 
                  subtitle = "Median sale price shown as a vertical red dashed line")

```

Across different binwidths, we can see the sales price data are right-skewed.  We can also see this by looking at a QQ-plot of the data on the original scale, and the data transformed on the log scale.

```{r}
ames %>% 
  select(Sale_Price) %>% 
  mutate(Log_Sale_Price = log10(Sale_Price)) %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "value") %>% 
  ggplot(aes(sample = value)) + 
  stat_qq(alpha = 0.2) + 
  stat_qq_line(linetype = "dashed") + 
  facet_wrap(~ type, scales = "free") + 
  labs(x = "Theoretical quantiles", 
       y = "Sample quantiles") +
  plot_annotation(title = "Log transformation reduces skew in sale price")
```

Since the log-transformation reduces skew in the data, we'll add a new predictor to log10 transform the sale price.

```{r}
# Apply log10 transformation
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

```

We'll use the `ames` data set with the log10 transformed sale price moving forward.


